---
title: "GitHub Actions Automate the Software Package Development Workflow"
author: |
  <br />
  Sean Davis, MD, PhD
date: October 30, 2019
output:
  revealjs::revealjs_presentation:
    self_contained: false
    theme: sky
    transition: zoom
    highlight: espresso
    reveal_plugins: ["notes", "zoom"]
vignette: >
  %\VignetteIndexEntry{GitHub Actions Automate the Software Package Development Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>"
)
```

```{r, setup}
library(BiocActions)
require(DiagrammeR)
```

# Bioconductor Workshop package automation

## Possible Workflow

```{r, grviz, echo=FALSE, fig.cap = "Potential workflow for a single package"}
DiagrammeR::grViz(diagram = 'bioc_example_workflow.dot')
```


## Multiple delivery targets for each workshop

- Docker container with all installed packages (or for each workshop
  independently).
- Website for each workshop
- Binary and source package tarballs for each workshop

## Broader goals align with developer interests

- Provide developers with access to a build environment as close to
  Bioconductor build environment as possible.
- Promote automated continuous integration and delivery (CI/CD) without
  building extensive infrastructure.
- Encourage sharing CI/CD best practices through code and documentation.
- Facilitate build pipelines that are reproducible, testable, and
  deliver all required build artifacts.


# Understanding GitHub Actions

## Actions

Actions are not useful by themselves. They are combined into workflows
which are then run.

Actions come in two flavors:

- Containerized
- Javascript

## File layout for an `Action` directory

```{sh eval=FALSE}
.
└── hello-world-docker-action
    ├── Dockerfile
    ├── README.md
    ├── action.yml
    └── entrypoint.sh
```

Translated into an R package `inst` directory, for example.

```{sh eval=FALSE}
inst
└── actions
    └── hello-world-docker-action
        ├── Dockerfile
        ├── README.md
        ├── action.yml
        └── entrypoint.sh
```

## A Dockerfile for containerized builds

```sh
# Container image that runs your code
FROM alpine:3.10

# Copies your code file from your action
#  repository to the filesystem path `/` of the container
COPY entrypoint.sh /entrypoint.sh

# Code file to execute when the
#  docker container starts up (`entrypoint.sh`)
ENTRYPOINT ["/entrypoint.sh"]
```

## Running a script inside docker

The `entrypoint.sh` file:

```sh
#!/bin/sh -l

echo "Hello $1"
time=$(date)
#
# set output--idiomatic to github actions
# 
echo ::set-output name=time::$time
```

- Inputs are handled in the usual "linux/unix" way. 
- Outputs represent variables that can be accessed by
  other steps in the workflow.



## Example action

```yaml
# action.yml
name: 'Hello World'
description: 'Greet someone and record the time'
inputs:
  who-to-greet:  # id of input
    description: 'Who to greet'
    required: true
    default: 'World'
outputs:
  time: # id of output
    description: 'The time we greeted you'
runs:
  using: 'docker'
  image: 'Dockerfile'
  args:
    - ${{ inputs.who-to-greet }}
```

# Actions for our use case

Focus on linux that can run inside bioconductor_full docker container.

1. checkout
2. `R CMD CHECK`
3. `R CMD BUILD` (source)
4. `R CMD BUILD` (binary on linux)
5. `R CMD INSTALL` 
6. `Rscript -e 'pkgdown::build_site'`
7. Push to gh-pages
8. Workshop-specific docker image
9. Capture build artifacts

## Checkout

actions/checkout

## Check package

R CMD Check....

## Build source package

R CMD build

## Build binary package
 
R CMD build --binary

## Install package

```{sh eval=FALSE}
R CMD INSTALL ${GITHUB_WORKSPACE} >& install.log
```

## Build pkgdown website

pkgdown::build_site

## Full example

```{sh eval=FALSE}
name: CICD

'on':
  - push
  - pull_request

jobs:
  build:
    runs-on: ubuntu-18.04
    container:
      image: bioconductor/bioconductor_full:devel
    steps:
      - name: Checkout Repository
        uses: actions/checkout@master
      - name: Install Package Dependencies
        run: |
          Rscript -e "BiocManager::install(c('remotes', 'rcmdcheck', 'covr', 'pkgdown'))"
          Rscript -e "remotes::install_deps(dependencies = TRUE)"
      - name: System dependencies
        run: |
          sudo apt-get update && apt-get install -y --no-install-recommends \
          texlive-latex-base texlive-fonts-recommended texlive-fonts-extra texlive-latex-extra
          pip install ghp-import
      - name: Check Package
        run: Rscript -e "rcmdcheck::rcmdcheck(error_on = 'error', check_dir = 'check')"
      - name: Document Package
        run: |
          Rscript -e "pkgdown::build_site()"
      - name: Deploy to GitHub Pages
        if: github.ref == 'refs/heads/master'
        run: |
          ghp-import -n docs && \
          git push -qf https://${GITHUB_TOKEN}@github.com/${GITHUB_REPOSITORY}.git gh-pages
```

## BiocActions

- https://github.com/seandavi/BiocActions


